### 2024-07-01 FinalThesis/Chapter_04/Manuscript.tex
Publish PHD

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage{scalerel}
\usepackage{makecell}
\usepackage{ifthen}
\usepackage{tikz}
\usepackage{lineno}
\usepackage{adjustbox}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%  Bibtex vs Biblatex
%------------------------------------------------------------------------------
% While bibtex is very nice to use, it's insanely slow. I therefore work with
% bibtex until the final stage. Set the value below to "true" or "false" to pick
% which one should be used.
\newboolean{usebiblatex}
\setboolean{usebiblatex}{true}

% Depending on the question, the commands below will prepare the document
% accordingly
\ifthenelse{\boolean{usebiblatex}}{
  %-----------------------------------------------------------------------------
  %  If using biblatex
  %-----------------------------------------------------------------------------
  % Load required packages and refer to correct bibliography
  \usepackage{csquotes}
  \usepackage[
      backend=biber
    , style=apa          % To display Author-year in the text
    , natbib=true        % Necessary to use citep, etc.
    , hyperref=true      % To have clickable links in the references
    , useeditor=false    % Don't add editor to publications
    , sorting=ynt        % Sort by year, name, and title
    , uniquename=false   % Avoid that initials are put
  ]{biblatex}
  \addbibresource{Literature.bib}

  % To include author in link
  \DeclareFieldFormat{citehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{#1}}
  \DeclareFieldFormat{textcitehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{%
    #1%
    \ifbool{cbx:parens}
      {\bibcloseparen\global\boolfalse{cbx:parens}}
      {}}}
  \savebibmacro{cite}
  \savebibmacro{textcite}
  \renewbibmacro*{cite}{%
    \printtext[citehyperref]{%
      \restorebibmacro{cite}%
      \usebibmacro{cite}}}
  \renewbibmacro*{textcite}{%
    \ifboolexpr{
      ( not test {\iffieldundef{prenote}} and
        test {\ifnumequal{\value{citecount}}{1}} )
      or
      ( not test {\iffieldundef{postnote}} and
        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
    }
      {\DeclareFieldAlias{textcitehyperref}{noformat}}
      {}%
    \printtext[textcitehyperref]{%
      \restorebibmacro{textcite}%
      \usebibmacro{textcite}}}

  % Remove editor form everything but books
  \AtEveryBibitem{%
   \ifentrytype{book}{}{%
    \clearname{editor}
   }
  }
}{
  %-----------------------------------------------------------------------------
  %  If using bibtex
  %-----------------------------------------------------------------------------
  % Load required packages and select citation style
  \usepackage[round]{natbib}
  \bibliographystyle{apalike}
}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Link colors
\definecolor{linkcolor}{HTML}{000000}
\hypersetup{
  colorlinks = true,
  linkcolor  = linkcolor,
  urlcolor   = linkcolor,
  citecolor  = linkcolor,
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% Avoid skip when including text from external .tex file
\newcommand{\inputy}[1]{\input{#1}\unskip}

% For supplementary material
\newcommand{\beginappendix}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
  \setcounter{equation}{0}
  \renewcommand{\theequation}{Equation S\arabic{equation}}%
  \setcounter{section}{0}
  \renewcommand{\thesection}{A.\arabic{section}}%
}

%------------------------------------------------------------------------------
%  ORCID
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Create the ORCID logo
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol]
      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
        ,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white]
      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

% Command to create the ORCID
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Methods for Implementing Integrated Step-Selection Functions with
Incomplete Data}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
  John Fieberg \textsuperscript{3} \orcid{0000-0002-3180-7021}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
Private Bag 13, Maun, Botswana.

\textsuperscript{3} Department of Fisheries, Wildlife, and Conservation Biology,
University of Minnesota, St. Paul, MN, USA.

\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}

\vspace{4cm}

\textbf{Running Title:} Step-Selection Analyses with Missing Data

\vspace{0.5cm}

\textbf{Keywords:} animal movement, gps data, imputation, incomplete data,
missing fixes, step-selection analyses, step-selection functions

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}

@@

\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
% \newpage
%
% \onehalfspacing
% \tableofcontents
% \doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

\section{Introduction}

@@

\begin{equation}
\label{EQ1}
u(s_{t+1}) = \frac{\phi(s_{t+1}, s_t, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)}{\int_{s \in G}\phi(s_{t+1}, s_{t}, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)ds}
\end{equation}

\noindent Here, the function $\phi$ represents an animal's \textbf{movement
kernel} which is usually expressed in terms of \textbf{step-length} and
\textbf{turning-angle} distributions, with $\gamma$ representing parameters in
these distributions. The function $w$ is the \textbf{habitat-selection function}
and reflects an animal's preferences $\beta$ for environmental characteristics
$x$ at location $s_{t+1}$. In most applications, $w$ is modeled as a log-linear
function of $x$, taking the form $w = exp(x^\top\beta)$. The integral in the
denominator of \ref{EQ1} ensures that $u$ is a proper probability distribution
(i.e., that it integrates to 1). Following \citet{Michelot.2024}, we call the
product $\phi \times w$ the step-selection function (SSF), as it highlights that
the probability of finding an animal at a certain location depends on both the
animal's movement kernel and its habitat-selection function.

Given a series of \textbf{observed steps}, finding the movement and
habitat-selection parameters that maximize the likelihood in \ref{EQ1} requires
approximating the integral in the denominator for each observed step. A variety
of numerical integration techniques can be used for this purpose
\citep{Michelot.2024}, but a common approach is to combine observed steps with
\textbf{random steps} generated by sampling step lengths and turning angles from
parametric distributions informed by the data \citep{Fortin.2005,
Thurfjell.2014}. Environmental conditions at observed steps are then contrasted
with environmental conditions at random steps in a (mixed effects) conditional
logistic regression framework \citep{Fortin.2005, Muff.2020}. To jointly
estimate parameters in $\phi$ and $w$, movement descriptors (e.g., step length
(sl), its natural logarithm (log(sl)), and the cosine of the turning angle
(cos(ta))) can be included in the conditional logistic regression model, and
their estimated coefficients can be used to update the initial (tentative)
step-length and turning-angle distributions \citep{Duchesne.2015, Avgar.2016,
Fieberg.2021}. The specific descriptors that need to be included depend on the
assumed step-length and turning-angle distributions \citep[for more details, see
Appendix C of][]{Fieberg.2021}. This approach to estimating parameters of the
SSF, termed \textit{integrated} SSA (or iSSA) by \citet{Avgar.2016}, is similar
to using importance sampling to approximate the integral in \ref{EQ1}
\citep{Michelot.2024} and is readily accessible through the \texttt{R}-package
\texttt{amt} \citep{Signer.2019}.

%In the SSF framework introduced by \citet{Fortin.2005}, the movement kernel
%$\phi$ was assumed to be known, which precluded joint estimation of $\phi$ and
%$w$ and has been shown to result in biased habitat-selection estimates
%\citep{Forester.2009}. A more robust alternative is to jointly estimate $\phi$
%and $w$ by fitting tentative distributions for step-lengths and turning-angles,
%and including descriptors of step-characteristics (i.e., step-lengths and
%turning-angles) as covariates in the statistical model \citep{Forester.2009,
%Avgar.2016}. In SSFs, the use of random steps serves to generate integration
%points that are needed to approximate the integral in \ref{EQ1} and, even
%though the true movement kernel $\phi$ is often not known a priori, it can be
%approximated using \textit{tentative} distributions \citep{Michelot.2024}. The
%use of parametric distributions serves as a means of conducting importance
%sampling and allows generating random steps that contribute better towards
%improving the approximation \citep{Michelot.2024}.

SSAs have proven extremely effective in numerous ecological studies
\citep{Thurfjell.2014}, providing insights into seasonal space use
\citep{Vales.2022, Enns.2023}, resource selection during distinct behavioral
phases \citep{Elliot.2014, Abrahms.2017, Cozzi.2018, Broekhuis.2019}, and the
effects of landscape familiarity or memory on animal movements \citep{Kim.2023}.
A model parametrized using iSSA resembles a fully mechanistic movement model
that can be used to simulate space use under novel conditions \citep{Avgar.2016,
Signer.2017, Hofmann.2023, Signer.2023}. This characteristic has made iSSAs a
useful tool for quantifying landscape resistance and identifying movement
corridors \citep{Zeller.2020, Buchholtz.2020, Hofmann.2021, Hofmann.2023}.

% Glossary of terms
\input{99_Glossary.tex}

@@

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Overview.png}
  \caption{(\textbf{a}) Demonstration of how missingness affects the number of
  valid steps that can be used for step-selection analyses under different
  levels of forgiveness. The upper panel depicts a trajectory with zero missing
  locations. That is, all aspired locations were successfully collected on a
  regular interval (yielding a regular step duration of $\Delta t = 1$). This
  trajectory produces four valid steps that can be included in the iSSF model
  and one invalid step that has to be omitted because it has no turning angle
  associated with it. In the central panel, animal location 4 was not obtained,
  introducing a missingness of 16.7\%. If the modeler has a forgiveness of one,
  only a single step can be included for further analysis, as all other steps
  are invalid (either because no turning angle can be computed or because
  step durations exceed the forgiveness). If, however, the modeler exhibited a
  forgiveness of two, such as in the lower panel, a total of three steps could
  be obtained for further analysis. (\textbf{b}) Conceptual illustration of how
  increasing the forgiveness allows one to retain additional steps that can be
  used for step-selection analysis. The sequence of dots resembles the sequence
  of locations that were scheduled to be collected (e.g. using a GPS device),
  with the lines representing hypothetical steps. Because not all locations were
  successfully obtained (gray dots), there is missingness. Depending on the
  forgiveness level, already a single missing location enforces the introduction
  of a new burst, which leads to the loss of several steps. In addition, some of
  the remaining steps are invalid (dotted) because they are lacking a
  turning angle. By increasing the forgiveness, a modeler is willing to retain
  steps that exceed the regular step duration by a certain threshold, which
  enables them to obtain longer bursts and increase the number of steps that can
  be used for further analysis. In the figure, forgiveness increases from left
  to right.}
  \label{Overview}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
  \caption{Illustration of how missingness in animal location data reduces the
  number of valid steps that can be used in step-selection analyses (left panel)
  and how increasing forgiveness helps to retain additional steps that are
  otherwise omitted (right panel). At a missingness of 0, 998 valid steps can be
  obtained from the total of 1,000 animal locations. At higher missingness,
  step durations become irregular, which means that the number of valid steps
  decreases substantially. However, if the modeler is willing to increase their
  forgiveness, additional steps can be gained. The right panel shows the number
  of valid steps that is gained when increasing the forgiveness from 1 to
  2, 3, 4, and 5, respectively. Ribbons indicate the 95\%-percentile intervals
  derived from 1,000 replicates.}
  \label{NumberOfSteps}
  \end{center}
\end{figure}

@@

  \item \textit{Imputation}: An intuitive solution is provided by
  \citet{McClintock.2017}, who suggests fitting a continuous-time correlated
  random walk movement model \citep{Johnson.2008} to the collected data and to
  use the fitted model to impute missing fixes. By imputing missing locations,
  the analysed trajectories become entirely regular again and can be analysed
  using traditional techniques. This approach, which we coin the
  \textit{imputation} approach, is readily available through the
  \texttt{R}-package \texttt{crawl} \citep{Johnson.2022}, yet has only been
  tested for use with hidden Markov movement models and not with iSSFs
  \citep{McClintock.2017}.

@@

\section{Methods}

We implemented the simulation study in the programming language \texttt{R}
version 4.3.2 \citep{RCoreTeam.2023} and achieved parallelization of
simulation-runs using the \texttt{R}-package \texttt{pbmcapply}
\citep{Kuang.2022}. We generated figures using the \texttt{ggplot2}
\citep{Wickham.2023}, \texttt{ggpubr} \citep{Kassambara.2023}, and
\texttt{ggh4x} \citep{Brand.2023} \texttt{R}-packages. We manipulated raster
data and computed spatial distances using the \texttt{R}-package \texttt{raster}
\citep{Hijmans.2023}. An overview of the simulation design and the different
iSSF approaches is presented in \Cref{Design} and all codes to reproduce this
study are available through an online repository \citep{Hofmann.2023a}.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Design.png}
  \caption{Illustration of the study design. We varied the autocorrelation range
  when simulating spatial covariates from 5 to 20 and tested for different
  missingness scenarios (ranging from 0\% to 50\% missing locations). To
  investigate how increasing forgiveness (i.e., the willingness to include steps
  with duration above the regular step duration) influenced model results, we
  varied its value from 1 (regular step selection) to 5 (considering steps that
  are five times the regular step duration). Finally, we tested five different
  methods to account for potential biases introduced by including irregular
  steps. This gave 3 x 6 x 5 x 5 = 450 combinations, each of which we replicated
  100 times. We assumed step lengths (sl) to follow a gamma distribution,
  whereas turning angles (ta) followed a von Mises distribution.}
  \label{Design}
  \end{center}
\end{figure}

@@

We simulated a virtual landscape comprising two continuous and one categorical
(binary) spatial layers, each with a resolution of 300 x 300 pixels
(\Cref{Covariates}) and spanning across x- and y-coordinates from 0 to 300. The
first layer, \textsf{Dist} (continuous), quantified the distance to the center
of the virtual landscape ($x = 150$, $y = 150$), and can be understood as a
point of attraction, such as, for instance, the center of an animal's
home-range. The second layer, \textsf{Elev} (continuous), resembled an elevation
layer and was simulated by sampling random pixel-values from a normal
distribution. To achieve spatial autocorrelation, we applied a circular moving
window with radius $r$ within which we tallied pixel-values. We varied $r$ from
5, to 10, to 20, depending on the simulated level of autocorrelation (Figure
S1). The third layer, \textsf{Forest} (categorical), represented areas covered
by woodland and was simulated similarly to the \textsf{Elev} layer, but we
binarized the layer by setting all simulated values above the 50\% quantile to
forest and all other values to non-forest (our reference class). We normalized
values of all simulated layers to a range between zero and one and replicated
the simulation of each layer 100 times per autocorrelation scenario, thus
producing 300 unique landscape configurations.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Covariates.png}
  \caption{Example of a landscape configuration across which we simulated
  movement trajectories. All simulated layers had a resolution of 300 x 300
  pixels. The distance layer indicated the distance to the center of the
  landscape and served to simulate attraction. The elevation and forest layers
  were simulated by sampling pixel-values from a normal distribution and
  applying a moving window to achieve spatial autocorrelation. Simulated
  individuals were initiated within the white dashed rectangle, which ensured
  that they would not be released directly at a map border. Simulated
  individuals were attracted to the landscape's center, preferred elevated
  areas, and avoided areas covered by forest. The black line shows the simulated
  trajectory associated with the visualized landscape configuration (cfr.
  \Cref{MovementSimulation}).}
  \label{Covariates}
  \end{center}
\end{figure}

@@

\begin{equation}
\label{EQ2}
  \pi_j =
  \frac{exp(\beta^\top x_j)}{\sum_{i = 1}^J exp(\beta^\top x_i)} %=
  % \frac{
  %   exp(\beta_{dist} dist_j + \beta_{elev} elev_j + \beta_{forest} forest_j)
  % }{
  %   \sum_{i = 1}^J exp(\beta_{dist} dist_i + \beta_{elev} elev_i + \beta_{forest} forest_i)
  % }
\end{equation}

@@

  \item \textit{Dynamic Distributions:} In this procedure, we fitted separate
  parametric distributions to step lengths and turning angles from steps of
  different step durations. That is, we parametrized separate turning-angle and
  step-length distributions representative of steps with durations of $\Delta t
  = 1, 2, 3, 4$ and $5$ (which corresponds to the maximum forgiveness level we
  tested for). Some step durations only rarely occurred at low levels of
  missingness, thus complicating parametrization of the associated
  distributions. To facilitate estimation of dynamic distribution parameters
  across all $\Delta t$ (Figure S2), we resampled data to different step
  durations using the \texttt{track\_resample} function from the \texttt{amt}
  package \citep{Signer.2019} before fitting tentative parameters. This ensured
  a sufficient number of steps for each step duration to estimate associated
  parameters. An alternative approach would be to increase missingness in the
  data even further, thus introducing a larger number of longer step durations.

>>>>

\documentclass[../FinalThesis.tex]{subfiles}
\begin{document}
\chapter{Methods for Implementing Integrated Step-Selection Functions with
Incomplete Data}
\label{ChapterIrregularity}
\thispagestyle{empty}
\vspace{-1cm}
\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil

\begin{center}
David D. Hofmann \orcid{0000-0003-3477-4365},
Gabriele Cozzi \orcid{0000-0002-1744-1940}, and
John Fieberg \orcid{0000-0002-3180-7021}

% \begin{flushleft}

% \vspace{5cm}
\vfill

Published in Movement Ecology (2024):
\url{https://doi.org/10.1186/s40462-024-00476-8}.

% \textsuperscript{1} Department of Evolutionary Biology and Environmental
% Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
% Switzerland.
%
% \vspace{0.25cm}
%
% \textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
% Private Bag 13, Maun, Botswana.
%
% \vspace{0.25cm}
%
% \textsuperscript{\S} Corresponding author:
% \href{mailto:david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}

\end{center}

% \vfill

% \textbf{Running Title:}\\African Wild Dog Dispersal in a Changing Climate:
% Lessons Learned from Seasonal Flood Extremes
%
% \vspace{0.5cm}
%
% \textbf{Keywords:}\\movement, connectivity, climate change, seasonality,
% dispersal, conservation, individual-based simulations, step-selection

% \end{flushleft}

\newpage
\section*{Abstract}

@@

\newpage
\section{Introduction}

@@

\begin{equation}
\label{EQ3}
u(s_{t+1}) = \frac{\phi(s_{t+1}, s_t, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)}{\int_{s \in G}\phi(s_{t+1}, s_{t}, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)ds}
\end{equation}

\noindent Here, the function $\phi$ represents an animal's \textbf{movement
kernel} which is usually expressed in terms of \textbf{step-length} and
\textbf{turning-angle} distributions, with $\gamma$ representing parameters in
these distributions. The function $w$ is the \textbf{habitat-selection function}
and reflects an animal's preferences $\beta$ for environmental characteristics
$x$ at location $s_{t+1}$. In most applications, $w$ is modeled as a log-linear
function of $x$, taking the form $w = exp(x^\top\beta)$. The integral in the
denominator of \ref{EQ3} ensures that $u$ is a proper probability distribution
(i.e., that it integrates to 1). Following \citet{Michelot.2024}, we call the
product $\phi \times w$ the step-selection function (SSF), as it highlights that
the probability of finding an animal at a certain location depends on both the
animal's movement kernel and its habitat-selection function.

% Glossary of terms
\input{Figures/Glossary.tex}

Given a series of \textbf{observed steps}, finding the movement and
habitat-selection parameters that maximize the likelihood in \ref{EQ3} requires
approximating the integral in the denominator for each observed step. A variety
of numerical integration techniques can be used for this purpose
\citep{Michelot.2024}, but a common approach is to combine observed steps with
\textbf{random steps} generated by sampling step lengths and turning angles from
parametric distributions informed by the data \citep{Fortin.2005,
Thurfjell.2014}. Environmental conditions at observed steps are then contrasted
with environmental conditions at random steps in a (mixed effects) conditional
logistic regression framework \citep{Fortin.2005, Muff.2020}. To jointly
estimate parameters in $\phi$ and $w$, movement descriptors (e.g., step length
(sl), its natural logarithm (log(sl)), and the cosine of the turning angle
(cos(ta))) can be included in the conditional logistic regression model, and
their estimated coefficients can be used to update the initial (tentative)
step-length and turning-angle distributions \citep{Duchesne.2015, Avgar.2016,
Fieberg.2021}. The specific descriptors that need to be included depend on the
assumed step-length and turning-angle distributions \citep[for more details, see
Appendix C of][]{Fieberg.2021}. This approach to estimating parameters of the
SSF, termed \textit{integrated} SSA (or iSSA) by \citet{Avgar.2016}, is similar
to using importance sampling to approximate the integral in \ref{EQ3}
\citep{Michelot.2024} and is readily accessible through the \texttt{R}-package
\texttt{amt} \citep{Signer.2019}.

%In the SSF framework introduced by \citet{Fortin.2005}, the movement kernel
%$\phi$ was assumed to be known, which precluded joint estimation of $\phi$ and
%$w$ and has been shown to result in biased habitat-selection estimates
%\citep{Forester.2009}. A more robust alternative is to jointly estimate $\phi$
%and $w$ by fitting tentative distributions for step-lengths and turning-angles,
%and including descriptors of step-characteristics (i.e., step-lengths and
%turning-angles) as covariates in the statistical model \citep{Forester.2009,
%Avgar.2016}. In SSFs, the use of random steps serves to generate integration
%points that are needed to approximate the integral in \ref{EQ3} and, even
%though the true movement kernel $\phi$ is often not known a priori, it can be
%approximated using \textit{tentative} distributions \citep{Michelot.2024}. The
%use of parametric distributions serves as a means of conducting importance
%sampling and allows generating random steps that contribute better towards
%improving the approximation \citep{Michelot.2024}.

SSAs have proven extremely effective in numerous ecological studies
\citep{Thurfjell.2014}, providing insights into seasonal space use
\citep{Vales.2022, Enns.2023}, resource selection during distinct behavioral
phases \citep{Elliot.2014, Abrahms.2017, Cozzi.2018, Broekhuis.2019}, and the
effects of landscape familiarity or memory on animal movements \citep{Kim.2023}.
A model parametrized using iSSA resembles a fully mechanistic movement model
that can be used to simulate space use under novel conditions \citep{Avgar.2016,
Signer.2017, Hofmann.2023, Signer.2024}. This characteristic has made iSSAs a
useful tool for quantifying landscape resistance and identifying movement
corridors \citep{Zeller.2020, Buchholtz.2020, Hofmann.2021, Hofmann.2023}.

@@

\begin{figure}[htpb]
  \begin{center}
  \includegraphics[width = \textwidth]{Figures/Overview.png}
  \caption{(a) Demonstration of how missingness affects the number of valid
  steps that can be used for step-selection analyses under different levels of
  forgiveness. The upper panel depicts a trajectory with zero missing locations.
  That is, all aspired locations were successfully collected on a regular
  interval (yielding a regular step duration of $\Delta t = 1$). This trajectory
  produces four valid steps that can be included in the iSSF model and one
  invalid step that has to be omitted because it has no turning angle associated
  with it. In the central panel, animal location 4 was not obtained, introducing
  a missingness of 16.7\%. If the modeler has a forgiveness of one, only a
  single step can be included for further analysis, as all other steps are
  invalid (either because no turning angle can be computed or because step
  durations exceed the forgiveness). If, however, the modeler exhibited a
  forgiveness of two, such as in the lower panel, a total of three steps could
  be obtained for further analysis. (b) Conceptual illustration of how
  increasing the forgiveness allows one to retain additional steps that can be
  used for step-selection analysis. The sequence of dots resembles the sequence
  of locations that were scheduled to be collected (e.g. using a GPS device),
  with the lines representing hypothetical steps. Because not all locations were
  successfully obtained (gray dots), there is missingness. Depending on the
  forgiveness level, already a single missing location enforces the introduction
  of a new burst, which leads to the loss of several steps. In addition, some of
  the remaining steps are invalid (dotted) because they are lacking a turning
  angle. By increasing the forgiveness, a modeler is willing to retain steps
  that exceed the regular step duration by a certain threshold, which enables
  them to obtain longer bursts and increase the number of steps that can be used
  for further analysis. In the figure, forgiveness increases from left to
  right.}
  \label{Overview}
  \end{center}
\end{figure}

\begin{figure}[htpb]
  \begin{center}
  \includegraphics[width = \textwidth]{Figures/NumberOfSteps.png}
  \caption{Illustration of how missingness in animal location data reduces the
  number of valid steps that can be used in step-selection analyses (left panel)
  and how increasing forgiveness helps to retain additional steps that are
  otherwise omitted (right panel). At a missingness of 0, 998 valid steps can be
  obtained from the total of 1,000 animal locations. At higher missingness,
  step durations become irregular, which means that the number of valid steps
  decreases substantially. However, if the modeler is willing to increase their
  forgiveness, additional steps can be gained. The right panel shows the number
  of valid steps that is gained when increasing the forgiveness from 1 to
  2, 3, 4, and 5, respectively. Ribbons indicate the 95\%-percentile intervals
  derived from 1,000 replicates.}
  \label{NumberOfSteps}
  \end{center}
\end{figure}

@@

  \item \textit{Imputation}: An intuitive solution is provided by
  \citet{McClintock.2017}, who suggests fitting a continuous-time correlated
  random walk movement model \citep{Johnson.2008} to the collected data and to
  use the fitted model to impute missing fixes. By imputing missing locations,
  the analysed trajectories become entirely regular again and can be analysed
  using traditional techniques. This approach, which we coin the
  \textit{imputation} approach, is readily available through the
  \texttt{R}-package \texttt{crawl} \citep{Johnson.2024}, yet has only been
  tested for use with hidden Markov movement models and not with iSSFs
  \citep{McClintock.2017}.

@@

\section{Materials and Methods}

We implemented the simulation study in the programming language \texttt{R}
version 4.3.2 \citep{RCoreTeam.2023} and achieved parallelization of
simulation-runs using the \texttt{R}-package \texttt{pbmcapply}
\citep{Kuang.2024}. We generated figures using the \texttt{ggplot2}
\citep{Wickham.2024}, \texttt{ggpubr} \citep{Kassambara.2024}, and
\texttt{ggh4x} \citep{Brand.2024} \texttt{R}-packages. We manipulated raster
data and computed spatial distances using the \texttt{R}-package \texttt{raster}
\citep{Hijmans.2024}. An overview of the simulation design and the different
iSSF approaches is presented in \Cref{Design} and all codes to reproduce this
study are available through an online repository \citep{Hofmann.2023a}.

\begin{figure}[htpb]
  \begin{center}
  \includegraphics[width = \textwidth]{Figures/Design.png}
  \caption{Illustration of the study design. We varied the autocorrelation range
  when simulating spatial covariates from 5 to 20 and tested for different
  missingness scenarios (ranging from 0\% to 50\% missing locations). To
  investigate how increasing forgiveness (i.e., the willingness to include steps
  with duration above the regular step duration) influenced model results, we
  varied its value from 1 (regular step selection) to 5 (considering steps that
  are five times the regular step duration). Finally, we tested five different
  methods to account for potential biases introduced by including irregular
  steps. This gave 3 x 6 x 5 x 5 = 450 combinations, each of which we replicated
  100 times. We assumed step lengths (sl) to follow a gamma distribution,
  whereas turning angles (ta) followed a von Mises distribution.}
  \label{Design}
  \end{center}
\end{figure}

@@

We simulated a virtual landscape comprising two continuous and one categorical
(binary) spatial layers, each with a resolution of 300 x 300 pixels
(\Cref{Covariates}) and spanning across x- and y-coordinates from 0 to 300. The
first layer, \textsf{Dist} (continuous), quantified the distance to the center
of the virtual landscape ($x = 150$, $y = 150$), and can be understood as a
point of attraction, such as, for instance, the center of an animal's
home-range. The second layer, \textsf{Elev} (continuous), resembled an elevation
layer and was simulated by sampling random pixel-values from a normal
distribution. To achieve spatial autocorrelation, we applied a circular moving
window with radius $r$ within which we tallied pixel-values. We varied $r$ from
5, to 10, to 20, depending on the simulated level of autocorrelation
(\Cref{SimulatedCovariates}). The third layer, \textsf{Forest} (categorical),
represented areas covered by woodland and was simulated similarly to the
\textsf{Elev} layer, but we binarized the layer by setting all simulated values
above the 50\% quantile to forest and all other values to non-forest (our
reference class). We normalized values of all simulated layers to a range
between zero and one and replicated the simulation of each layer 100 times per
autocorrelation scenario, thus producing 300 unique landscape configurations.

\begin{figure}[htpb]
  \begin{center}
  \includegraphics[width = \textwidth]{Figures/Covariates.png}
  \caption{Example of a landscape configuration across which we simulated
  movement trajectories. All simulated layers had a resolution of 300 x 300
  pixels. The distance layer indicated the distance to the center of the
  landscape and served to simulate attraction. The elevation and forest layers
  were simulated by sampling pixel-values from a normal distribution and
  applying a moving window to achieve spatial autocorrelation. Simulated
  individuals were initiated within the white dashed rectangle, which ensured
  that they would not be released directly at a map border. Simulated
  individuals were attracted to the landscape's center, preferred elevated
  areas, and avoided areas covered by forest. The black line shows the simulated
  trajectory associated with the visualized landscape configuration (cfr.
  \Cref{MovementSimulation}).}
  \label{Covariates}
  \end{center}
\end{figure}

@@

\begin{equation}
\label{EQ4}
  \pi_j =
  \frac{exp(\beta^\top x_j)}{\sum_{i = 1}^J exp(\beta^\top x_i)} %=
  % \frac{
  %   exp(\beta_{dist} dist_j + \beta_{elev} elev_j + \beta_{forest} forest_j)
  % }{
  %   \sum_{i = 1}^J exp(\beta_{dist} dist_i + \beta_{elev} elev_i + \beta_{forest} forest_i)
  % }
\end{equation}

@@

  \item \textit{Dynamic Distributions:} In this procedure, we fitted separate
  parametric distributions to step lengths and turning angles from steps of
  different step durations. That is, we parametrized separate turning-angle and
  step-length distributions representative of steps with durations of $\Delta t
  = 1, 2, 3, 4$ and $5$ (which corresponds to the maximum forgiveness level we
  tested for). Some step durations only rarely occurred at low levels of
  missingness, thus complicating parametrization of the associated
  distributions. To facilitate estimation of dynamic distribution parameters
  across all $\Delta t$ (\Cref{DistributionParameters}), we resampled data to
  different step durations using the \texttt{track\_resample} function from the
  \texttt{amt} package \citep{Signer.2019} before fitting tentative parameters.
  This ensured a sufficient number of steps for each step duration to estimate
  associated parameters. An alternative approach would be to increase
  missingness in the data even further, thus introducing a larger number of
  longer step durations.
